---
title: "Circle plots"
author: "Alexandria Andrayas"
date: "13/09/2023"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE, tidy=T)
```

```{r, echo=FALSE, message=FALSE}
source('C:/Users/qg21962/OneDrive - University of Bristol/Documents/CRUK smoking vaping/Longitudinal smoking/Scripts/circ_plts_html.R')
```

## Non-smoking vs any smoking
    
### Family substance use

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial substance use'),], c(1:4))
```

### Family sociodemographics   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial sociodemographics'),], c(1:4))
```

### Sociodemographic factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Sociodemographic factors'),], c(1:4))
```

### Lifestyle factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Lifestyle factors'),], c(1:4))
```

### Mental health and other factors

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Mental health and other factors'),], c(1:4))
```


## Occasional smoking vs any regular smoking

### Family substance use

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial substance use'),], c(5:7))
```

### Family sociodemographics   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial sociodemographics'),], c(5:7))
```

### Sociodemographic factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Sociodemographic factors'),], c(5:7))
```

### Lifestyle factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Lifestyle factors'),], c(5:7))
```

### Mental health and other factors

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Mental health and other factors'),], c(5:7))
```

## Short-term smoking vs more persistent smoking

### Family substance use

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial substance use'),], c(8:9))
```

### Family sociodemographics   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial sociodemographics'),], c(8:9))
```

### Sociodemographic factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Sociodemographic factors'),], c(8:9))
```

### Lifestyle factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Lifestyle factors'),], c(8:9))
```

### Mental health and other factors

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Mental health and other factors'),], c(8:9))
```

## Late vs early-onset smoking  

### Family substance use

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial substance use'),], 10)
```

### Family sociodemographics   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Familial sociodemographics'),], 10)
```

### Sociodemographic factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Sociodemographic factors'),], 10)
```

### Lifestyle factors   

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Lifestyle factors'),], 10)
```

### Mental health and other factors

```{r, out.width="100%", out.height="100%", dpi=300, echo=FALSE, message=FALSE, results='asis'}
circlplt(allests[allests$lab_grp %in% c('Mental health and other factors'),], 10)
```
