---
title: "Longitudinal patterns of smoking behaviours in adolescence and early adulthood and their association with modifiable and sociodemographic risk factors - Interactive plots"
author: "Andrayas et al"
date: "2023"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load-packages, include=FALSE}
library(htmlwidgets)
library(ggplot2)
library(plotly)
```

```{r, echo=FALSE, message=FALSE}
source('C:/Users/qg21962/OneDrive - University of Bristol/Documents/CRUK smoking vaping/Longitudinal smoking/Scripts/int_plts_all_html.R')
load("C:/Users/qg21962/OneDrive - University of Bristol/Documents/CRUK smoking vaping/Longitudinal smoking/Figures/int_plts_all.Rda")
```

## Comparison  
<br>

### Non-smoking vs
<br>

#### Early-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C2VC1
```

#### Late-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C5VC1
```

#### Short-term smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C3VC1
```

#### Occasional smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C4VC1
```

### Occasional smoking vs
<br>

#### Early-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C2VC4
```

#### Late-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C5VC4
```

#### Short-term smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C3VC4
```

### Short-term smoking vs
<br>

#### Early-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C2VC3
```

#### Late-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C5VC3
```

### Late-onset smoking vs
<br>

#### Early-onset smoking

```{r, fig.width=3, fig.height=2, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$C2VC5
```

## Theme
<br>

### Family and peer smoking

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$fam_peer_smk
```

### Family and peer substance use

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$fam_peer_sub
```

### Family sociodemographic factors

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$fam_socdem
```

### Individual lifestyle factors

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$lifestyle
```

### Individual sociodemographic factors

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$socdem
```

### Mental health and other factors

```{r, fig.width=3, fig.height=4, dpi=300, echo=FALSE, message=FALSE, results='asis'}
p$mhoth
```
